Trauma treatment delay

### 2023-09-19 manuscriptlongversion.Rmd
intro and code

Once treated, the patients information and timeline of what happened enters a trauma care database, depending on what hospital. At Nya Karolinska Universitetssjukhuset, all patients who have been treated and had ISS\>9 or had the trauma alarm activated are included and screened for OFIs. A cornerstone of trauma quality improvement programs is multidisciplinary mortality and morbidity review to identify OFIs and implement corrective actions. This refers to a meeting where doctors from different specialities sit down and go through cases where possible OFIs were identified by nurses who go through the trauma care quality database. Examples of opportunities for improvement may include lack of resources and management errors. One common OFI that this paper will look further into is delayed treatment. Among the preventable or possibly preventable deaths in trauma patients, a subgroup of patients determined by the Combined Trauma Death Review Committee, delay in treatment has been identified as a major error contributing to death, found in up to 52.9% of patients in said group [@Teixeira_Inaba_Hadjizacharia_Brown_Salim_Rhee_Browder_Noguchi_Demetriades_2007]. Although it is such a common OFI, the patient level factors associated with delayed treatment are poorly understood.

@@

This is a registry based cohort study.

@@

The data used is from the trauma registry and trauma care quality database at the Karolinska University Hospital in Solna. The trauma registry includes about 12000 patients treated between 2012 and 2022. The trauma care quality database is a subset of the trauma registry and includes about 6000 patients selected for review between 2014 and 2022.

>>>>

Once treated, the patients information and timeline of what happened enters a trauma care database, depending on what hospital. At Nya Karolinska Universitetssjukhuset, all patients who have been treated and had ISS\>9 or had the trauma alarm activated are included and screened for OFIs. A cornerstone of trauma quality improvement programs is multidisciplinary mortality and morbidity review to identify OFIs and implement corrective actions. This refers to a meeting where doctors from different specialities sit down and go through cases where possible OFIs were identified by nurses who go through the trauma care quality database. Examples of opportunities for improvement may include lack of resources and management errors. One common OFI that this paper will look further into is delayed treatment. Among the preventable or possibly preventable deaths in trauma patients, a subgroup of patients determined by the Combined Trauma Death Review Committee, delay in treatment has been identified as a major error contributing to death, found in up to 52.9% of patients in said group [@Teixeira_Inaba_Hadjizacharia_Brown_Salim_Rhee_Browder_Noguchi_Demetriades_2007]. Although it is such a common OFI, the patient level factors associated with delayed treatment are poorly understood. There may be several factors that correlate with receiving delayed treatment, but they have yet to be identified. Mapping these factors may help in identifying patients that might be at risk for receiving delayed treatment before it happens.

@@

This is a registry based cohort study using data from the trauma registry and trauma care quality database at the Karolinska University Hospital in Solna. The trauma registry includes about 12000 patients treated between 2012 and 2022. The trauma care quality database is a subset of the trauma registry and includes about 6000 patients selected for review between 2014 and 2022. The outcome is delayed treatment, as identified by the multidisciplinary review board and recorded in the trauma care quality database. This project will link the two databases and assess how different patient level factors, such as age, sex, mechanism of injury, and injury severity, are associated with delayed treatment using logistic regression. A 5% significance level and 95% confidence levels will be used.

@@

Every year, Karolinska University hospital, which is the equivalent of a trauma level 1 hospital, admits 1500 trauma patients each year to their facility. To be added to the trauma registry, a patient must be over 15 years of age, had ISS\>9 and/or had trauma code activation. They get added to both the Karolinska Trauma registry but also the national trauma registry (SweTrau). The registry includes data on vital signs, injuries, interventions and patient information. A subset to this database, he trauma care quality database includes information relevant for mortality and morbidity conferences
